Under-approximation of reachability in multivalued asynchronous networks

### 2013-04-02 parts/ccl.tex
proofreading of conclusion

\section{Discussion \& Conclusion}\label{sec:ccl}
\towrite{Pistes :
\begin{itemize}
  \item Plus de classes de priorités ?
  \item Modèles Zeno (système de précédences et prise en compte des états globaux)
\end{itemize}
}

>>>>

\section{Discussion \& Conclusion}\label{sec:ccl}

We introduced a new semantics to include priorities into the Process Hitting framework, which prove especially useful to model cooperations.
Then, we developed a method to efficiently perform a reachability analysis of a sequence of objectives in a restricted class of Process Hitting models,
but it is also useful to establish the reachability of a partial state.
This analysis is based on an under-approximation of the true reachability solutions.
%; however, the most usual cases can be handled.
The method has been implemented as an extension of the existing “ph-reach” tool into the existing Pint\footnote{\url{http://processhitting.wordpress.com/}} library.
\towrite{Déplacer dans la section exemples ?}

We showed that the class of Process Hitting models that can be handled by the aforementioned method are equivalent to Asynchronous Discrete Networks, and therefore to Asynchronous Boolean Networks.
This allows to efficiently compute reachability results on large biological models provided that they are equivalent to Asynchronous Discrete Networks and that a translation from the original framework into a Process Hitting model is possible.
Such a translation for interaction graphs of Thomas modelling was proposed in~\cite{PMR10-TCSB} \towrite{[À garder ?] and is made possible by the Pint software}.

Further work can be derived from what have been presented in this paper.
The over-approximation on Process Hitting models without priorities proposed in~\cite{PMR12-MSCS}
is still accurate in the framework with priorities (by “flattening” all actions),
but may be refined given the restrictions proposed in this paper.
%but turns out to be too wide even in some obvious cases that are consequently not conclusive.
%This approximation may be refined in order to better fit the introduction of priorities, and mak the overall approximation approach 
%and mode precisely the class of models studied in this paper.
Furthermore, a more general under-approximation could be developed in order to handle a larger class of Process Hitting models, that is,
models with more than two classes of priorities, that do not only contain components of cooperative sorts, or whose behaviour may contain cycles or cyclic attractors.
Finally, in order to take into account quantitative data in transition delays, the overall approximation method could be extended to handle evolution that are chronometric instead of only chronologic.

### 2013-04-02 parts/def-ph.tex
proofreading of section 3
proofreading of conclusion

\begin{figure}
  \centering
  \begin{tikzpicture}
    \TSort{(0,0)}{a}{2}{l}
    \TSort{(0,4)}{b}{2}{l}
    \TSort{(6,2)}{c}{2}{r}

@@

    \path[bounce, bend right]
      \TBounce{ab_3}{}{ab_1}{.north west}
      \TBounce{ab_2}{}{ab_0}{.north west}
      \TBounce{ab_3}{}{ab_2}{.north west}
      \TBounce{ab_1}{}{ab_0}{.north west}
    ;
    \path[bounce, bend left]
      \TBounce{ab_1}{}{ab_3}{.south west}
      \TBounce{ab_0}{}{ab_2}{.south west}
      \TBounce{ab_2}{}{ab_3}{.south west}
      \TBounce{ab_0}{}{ab_1}{.south west}
    ;
    \path[bounce, bend right]
      \TBounce{a_1}{}{a_0}{.north west}
      \TBounce{b_1}{}{b_0}{.north west}
    ;
    \path[bounce, bend left]
      \TBounce{a_0}{}{a_1}{.south west}
      \TBounce{b_0}{}{b_1}{.south west}
    ;
    \path[bounce, bend left]
      \TBounce{c_0}{}{c_1}{.south west}
    ;
    \TState{a_1, b_0, ab_2, c_0}
  \end{tikzpicture}
  \caption{
  \label{fig:ph-livelock}
    An example of PH with $2$ classes of priorities.
    Sorts are represented as labelled boxes and processes as circles with their identifier on the side.
    Actions of $\PHh^{(1)}$ are represented by double arrows and actions of $\PHh^{(2)}$ are represented by single arrows;
    the hit part of each action in drawn in plain line and the bounce part is in dotted line.
    Greyed processes stand for a possible state $s = \PHstate{a_1, b_0, c_0, ab_{10}}$.
    %This PH notably contains three well-formed components ($a$, $b$ and $c$) and a well-formed cooperative sort ($ab$).
  }
\end{figure}

@@

Finally, we introduce the notion of local steady-state of a sort from a given state.
This local steady-state is the set of processes towards which the sort tends to evolve to, and stay in, when playing only prioritised actions.
We denote $\pfp_s(a)$ this set of processes for a sort $a$ in state $s$ (\pref{def:pfp}),
and we derive from this the notion of well-formed cooperative sort (\pref{def:cs}).
A well-formed cooperative sort models a cooperation between sorts as presented in \pref{ssec:cooperation};
therefore, it must be only hit by prioritised actions,
in a way that its local steady-states efficiently represents all configurations of its neighbouring sorts.
%any sub-state of its neighbourhood is represented by exactly one local steady state in this sort,
%and all of its processes represents at least one sub-state of its neighbourhood. \towrite{reformuler ou supprimer}
\begin{definition}[$\pfp : \PHl \times \PHs \rightarrow \powerset(\PHproc)$]
\label{def:pfp}
  For all $s \in \PHl$ and $a \in \PHs$,
  \begin{align*}
    \pfp_s(a) = \{ a_i \in \PHl_a &\mid \exists \delta \in \restriction{\Sce}{1}(s), \PHget{(s \PHplay \delta)}{a} = a_i \\
    & \quad \wedge \forall \PHhit{b_i}{c_j}{c_k} \in \Vh(a), \PHget{(s \PHplay \delta)}{b} \neq b_i \vee \PHget{(s \PHplay \delta)}{c} \neq c_j \}
  \end{align*}
  Of course, if $a \in \components$, then $\pfp_s(a) = \{ \PHget{s}{a} \}$.
\end{definition}
%
\begin{definition}[Well-formed cooperative sorts ($\cs$)]
\label{def:cs}
  A sort $a \in \PHs$ is a \emph{well-formed cooperative} sort if and only if:
  \begin{enumerate}
    \item $\exists b \in \PHs, \PHh(b, a) \neq \emptyset$
%    \item $\forall s \in \underset{b \in \Vs(a)}{\times} \PHl_b, |\pfp_s(a)| = 1$
%    \item $\forall a_i \in \PHl_a, \exists s \in \underset{b \in \Vs(a)}{\times} \PHl_b, \pfp_s(a) = \{ a_i \}$
    \item \label{csai} $\forall s \in \PHl, \exists a_i \in \PHl_a, \pfp_s(a) = \{ a_i \}$
    \item \label{css} $\forall a_i \in \PHl_a, \exists s \in \PHl, \pfp_s(a) = \{ a_i \}$
  \end{enumerate}
  We call $\cs$ the set of well-formed cooperative sorts.
\end{definition}

@@

The criteria given in this subsection allow to define a class of PH models that is (weakly) bisimilar to Asynchronous Boolean Networks
or, more generally, to the multivalued version of this framework, called Asynchronous Discrete Models (ADM).
A translation of ADM into PH is given in \pref{suppl:demotrad}, alongside with a demonstration of the weak bisimulation.

>>>>

\begin{figure}[p]
  \centering
  \scalebox{1.5}{
  \begin{tikzpicture}
    \TSort{(0,0)}{a}{2}{l}
    \TSort{(0,4)}{b}{2}{l}
    \TSort{(6,2)}{c}{2}{r}

@@

    \path[bounce, bend right]
      \TBounce{ab_3}{}{ab_1}{.north west}
      \TBounce{ab_2}{}{ab_0}{.north west}
      \TBounce{ab_3}{}{ab_2}{.north west}
      \TBounce{ab_1}{}{ab_0}{.north west}
    ;
    \path[bounce, bend left]
      \TBounce{ab_1}{}{ab_3}{.south west}
      \TBounce{ab_0}{}{ab_2}{.south west}
      \TBounce{ab_2}{}{ab_3}{.south west}
      \TBounce{ab_0}{}{ab_1}{.south west}
    ;
    \path[bounce, bend right]
      \TBounce{a_1}{}{a_0}{.north west}
      \TBounce{b_1}{}{b_0}{.north west}
    ;
    \path[bounce, bend left]
      \TBounce{a_0}{}{a_1}{.south west}
      \TBounce{b_0}{}{b_1}{.south west}
    ;
    \path[bounce, bend left]
      \TBounce{c_0}{}{c_1}{.south west}
    ;
    \TState{a_1, b_0, ab_2, c_0}
  \end{tikzpicture}
  }
  \caption{
  \label{fig:ph-livelock}
    An example of PH with $2$ classes of priorities.
    Sorts are represented as labelled boxes and processes as circles with their identifier on the side.
    Actions of $\PHh^{(1)}$ are represented by double arrows and actions of $\PHh^{(2)}$ are represented by single arrows;
    the hit part of each action in drawn in plain line and the bounce part is in dotted line.
    Greyed processes stand for a possible state $s = \PHstate{a_1, b_0, c_0, ab_{10}}$.
    %This PH notably contains three well-formed components ($a$, $b$ and $c$) and a well-formed cooperative sort ($ab$).
  }
\end{figure}

@@

Finally, we introduce the notion of local steady-state of a sort from a given state.
This local steady-state is the set of processes towards which the sort tends to evolve to, and stay in, when playing only prioritised actions.
We denote $\pfp_s(a)$ this set of processes for a sort $a$ in state $s$ (\pref{def:pfp}),
and we derive from this the notion of well-formed cooperative sort (\pref{def:cs}).
A well-formed cooperative sort models a cooperation between sorts as presented in \pref{ssec:cooperation};
therefore, it must be only hit by prioritised actions,
in a way that its local steady-states efficiently represents all configurations of its neighbouring sorts.
%any sub-state of its neighbourhood is represented by exactly one local steady state in this sort,
%and all of its processes represents at least one sub-state of its neighbourhood. \towrite{reformuler ou supprimer}
\begin{definition}[$\pfp : \PHs \rightarrow \powerset(\PHproc)$]
\label{def:pfp}
  For all $a \in \PHs$,
  \begin{align*}
    \pfp_s(a) = \{ a_i \in \PHl_a &\mid \exists \delta \in \restriction{\Sce}{1}(s), \PHget{(s \PHplay \delta)}{a} = a_i \\
    & \quad \wedge \forall \PHhit{b_i}{c_j}{c_k} \in \Vh(a), \PHget{(s \PHplay \delta)}{b} \neq b_i \vee \PHget{(s \PHplay \delta)}{c} \neq c_j \}
  \end{align*}
  Of course, if $a \in \components$, then $\pfp_s(a) = \{ \PHget{s}{a} \}$.
\end{definition}
%
\begin{definition}[Well-formed cooperative sorts ($\cs$)]
\label{def:cs}
  A sort $a \in \PHs$ is a \emph{well-formed cooperative} sort if and only if:
  \begin{enumerate}
    \item $\exists b \in \PHs, \PHh(b, a) \neq \emptyset$
%    \item $\forall s \in \underset{b \in \Vs(a)}{\times} \PHl_b, |\pfp_s(a)| = 1$
%    \item $\forall a_i \in \PHl_a, \exists s \in \underset{b \in \Vs(a)}{\times} \PHl_b, \pfp_s(a) = \{ a_i \}$
    \item \label{csai} $\forall s \in \PHl, \exists a_i \in \PHl_a, \pfp_s(a) = \{ a_i \}$
    \item \label{css} $\forall a_i \in \PHl_a, \exists s \in \PHl, \pfp_s(a) = \{ a_i \}$
  \end{enumerate}
  We call $\cs$ the set of well-formed cooperative sorts.
\end{definition}

@@

\begin{example}
  The PH in \pref{fig:ph-livelock} contains three components $a$, $b$ and $c$ and a cooperative sort $ab$ that models cooperation between sorts $a$ and $b$.
\end{example}

The criteria given in this subsection allow to define a class of PH models that is (weakly) bisimilar to Asynchronous Boolean Networks
or, more generally, to the multivalued version of this framework, called Asynchronous Discrete Models (ADN).
A translation of ADN into PH is given in \pref{suppl:demotrad}, alongside with a demonstration of the weak bisimulation.
